---
title: "qstack and dequer"
output: html_document
---

```{r, echo=FALSE}
library(qstack)
library(dequer)

library(knitr)
opts_chunk$set(cache=TRUE)
```


## Many small objects

Winner:  qstack

```{r}
n <- 2e5
index <- 1:n


system.time({
  q <- Queue2()
  for (i in index) q$add(i)
})

system.time({
  d <- deque()
  
  for (i in index) push(d, i)
})

invisible({rm(q, d);gc()})
```



## Copies of the same object

Winner:  overwhelmingly qstack (it's using a hash table, after all)

```{r}
n <- 1e3
index <- 1:n

x <- rnorm(1e5)


system.time({
  q <- Queue2()
  for (i in index) q$add(x)
})

system.time({
  d <- deque()
  
  for (i in index) push(d, x)
})

invisible({rm(q, d);gc()})
```



## Sevaral larger objects

Winner: dequer

```{r}
n <- 1e4
index <- 1:n

m <- 1e3


system.time({
  q <- Queue2()
  for (i in index) q$add(rnorm(m))
})

system.time({
  d <- deque()
  
  for (i in index) push(d, rnorm(m))
})

invisible({rm(q, d);gc()})
```

I first tried this with `1e3` iterations with `rnorm(1e6)`, and it made my 16 GiB ram machine swap out from the Queue2 insertions.



## Thoughts

* If the memory allocations are expensive, dequer (linked list) is probably going to be faster than qstack (doubling array), and otherwise I would would expect qstack to be faster than dequer.
* These benchmarks basically follow the general wisdom on queue-like data structures.
* As documented in the dequer package, R's memory allocator, which dequer interfaces with via `R_PreserveObject` and `R_ReleaseObject`, is a stack.  This makes the full generality of dequer somewhat worthless.  Observe:

```{r}
n <- 1e4
index <- 1:n

system.time({
	d <- deque()
	for (i in index) pushback(d, i)
	
	invisible({rm(d);gc()})
})

### Reverse first to free the stack in the correct order
system.time({
	d <- deque()
	for (i in index) pushback(d, i)
	rev(d)
	
	invisible({rm(d);gc()})
})
```

If you limit yourself to just pushes or pushbacks (stacks or queues), you have the extra amount of information needed to ensure that you always release "correctly" and avoid this penalty.